PAPER DRAFT

### 2021-11-10 Papers/Polymethylation_scores/01_titlepage.Rmd
kelly revisions round 2

@@

**Authors:** Freida A. Blostein^1, Jonah Fisher^2, John Dou^1, Lisa Schenper^3, Erin Ware^2 Daniel A. Notterman^3, Colter Mitchell^2, Kelly M. Bakulski^1

@@

^1 Department of Epidemiology, University of Michigan School of Public Health, Ann Arbor, Michigan, United States of America

^2 Institute for Social Research, University of Michigan, Ann Arbor, Michigan, United States of America

^3 Department of Molecular Biology, Princeton University, Princeton, New Jersey, United States of America

>>>>

---
output:
  html_document: default
  word_document: default
bibliography: Methylation_SmokingBib.bib
always_allow_html: true
---

@@

**Authors:** Freida A. Blostein$^1$, Jonah Fisher$^2$,  John Dou$^1$, Lisa Schenper$^3$, Erin Ware$^2$ Daniel A. Notterman$^3$, Colter Mitchell$^2$, Kelly M. Bakulski$^1$`

@@

$^1$Department of Epidemiology, University of Michigan School of Public Health, Ann Arbor, Michigan, United States of America

$^2$Institute for Social Research, University of Michigan, Ann Arbor, Michigan, United States of America

$^3$Department of Molecular Biology, Princeton University, Princeton, New Jersey, United States of America

### 2021-11-10 Papers/Polymethylation_scores/08_methods.Rmd
kelly revisions round 2

#### Primary model

In our primary model we adjusted for maternal income to poverty ratio at baseline, child sex, child age, plate from DNA methylation processing, and estimated immune cell proportion estimated from DNA methylation. In nonstratified models, we adjusted for the first two components of genetic ancestry from principal component analysis. In ancestry-stratified models, the first two principal components from principal component analysis run within each ancestry strata were used. While child sex, child age, plate from DNA methylation processing and immune cell proportions are not confounders (as they cannot casually effect prenatal maternal smoke exposure), these variables can strongly effect DNA methylation and so were adjusted for as precision variables.

>>>>

#### Base model

In our base model we adjusted for maternal income to poverty ratio at baseline, child sex, child age, plate from DNA methylation processing, and estimated immune cell proportion estimated from DNA methylation. In nonstratified models, we adjusted for the first two components of genetic ancestry from principal component analysis. In ancestry-stratified models, the first two principal components from principal component analysis run within each ancestry strata were used. While child sex, child age, plate from DNA methylation processing and immune cell proportions are not confounders (as they cannot casually effect prenatal maternal smoke exposure), these variables can strongly effect DNA methylation and so were adjusted for as precision variables.
